Recurrent event analysis in the presence of real-time high frequency data via random subsampling

### 2020-08-01 write-up/fda-recurrent.tex
Updated paper for simulation results

We next assess the proposed methodology via a simulation study. Here, we assume each individual is observed over five days where each day is defined on the unit interval~$[0,1]$ with $1000$ equally spaced observation times per day. We define~$X(t)$ at the grid of observations as a mean-zero Gaussian process with covariance
\[
  \Sigma (t, t^\prime) =
  \frac{\sigma^2}{\Gamma (v) 2^{\nu-1}}
  \left( \frac{\sqrt{2 \nu} |t-s|}{\rho_t} \right)^{\nu}
  K_v \left( \frac{\sqrt{2 \nu} |t-s|}{\rho_t} \right)
\]
where~$K_v$ is the modified Bessel function of the second kind.  We set~$\nu = 1/2$, $\sigma^2 = 1$, and~$\rho = 0.3$ as well as set~$K_b = K_x = 35$.  For simplicity, we assume~$\Sigma$ is known in computation of the eigendecompositions. Given~$\{ X(t) \}_{0 \leq t \leq 1}$ for a given user-day, we generate event times according a chosen hazard function~$h (t; \theta)$.  To mimic our real data, we set
\[
h(t; \theta) = \exp \left( \theta_0 + \int_{0}^{\Delta} X(t-s) \beta(s) ds \right).
\]
We set~$\Delta = 1/(12*2)$ to mimic a $30$-minute window for a $12$-hour day.  We set~$\theta_0 = \log(5/1000)$ to set a baseline risk of approximately 5 events per day. We consider two choices of~$\beta(s)$: (1) $\beta_0 + \exp(- \beta_1 s) $, and (2) $\beta_1  * \sin \left( 2 \pi \frac{s}{\Delta} - \pi/2 \right)$

\begin{itemize}
\item Sample every 0.5 hours; then subsample at each subrate. Estimates are given by $\hat \beta_{i,0.5} (t)$ and $\hat \beta_{i,1} (t)$.
\item Bias is average over the estimates $\bar \hat \beta_{0.5} (t) = 100^{-1} \sum_{i=1}^100 \hat \beta_{i, 0.5} (t)$ compared to truth.
\item Take squared difference integrated to get variance.
\item To get the subsampling variance take $1/1000 \sum_{i=1}^1000 (\hat \beta_{i,0.5} (t) - \hat \beta_{i,1} (t))^2$.
\item Normalize everything by squared function!
\end{itemize}

We sample non-events at rates of 1 per 12/6/1/0.5 hours.

Since we are primarily interested in accuracy of the final estimates~$\hat \beta(s)$, we report the mean integrated squared error (MISE) of~$\hat \beta (s)$, defined as~$\frac{1}{100} \sum_{j=1}^{100} \int_{0}^{\Delta^\prime} ( \hat \beta_j (s) - \beta(s))^2 ds$ where~$\beta(s) \equiv 0$ for $s > \Delta$.  We define the MISE in this manner to account for settings where~$\Delta$ is unknown. Table~\ref{tab:mise} show the MISE decomposed into the variance and squared biase

\begin{table}[!th]
\begin{tabular}{c | c c c c c}
& \multicolumn{5}{c}{Sampling rate} \\ \cline{2-6}
 & 0.5 & 1 & 5 & 25 & 50 \\ \hline
Subsampling variance & (10,12) & - & - & - & - \\
Variance & (10,12) & - & - & - & - \\
Squared Bias & - & - & - & - & -  \\
MISE & - & - & - & - & -  \\ \hline
Type I error & - & - & - & - & -  \\ \hline
Avg runtime & - & - & - & - & - \\
\end{tabular}
\caption{Mean-integrated squared error and type I error as a function of sampling rate for $\beta(s)$ given by (1) and (2) respectively.}
\label{tab:mise}
\end{table}

\subsection{Coverage properties}

We simulate the system $1000$ times under $\beta \equiv 0$ and check the type I error.

@@

>>>>

We next assess the proposed methodology via a simulation study. Here, we assume each individual is observed over five days where each day is defined on the unit interval~$[0,1]$ with $1000$ equally spaced observation times per day. We define~$X(t)$ at the grid of observations as a mean-zero Gaussian process with covariance
\[
  \Sigma (t, t^\prime) =
  \frac{\sigma^2}{\Gamma (v) 2^{\nu-1}}
  \left( \frac{\sqrt{2 \nu} |t-s|}{\rho_t} \right)^{\nu}
  K_v \left( \frac{\sqrt{2 \nu} |t-s|}{\rho_t} \right)
\]
where~$K_v$ is the modified Bessel function of the second kind.  We set~$\nu = 1/2$, $\sigma^2 = 1$, and~$\rho = 0.3$ as well as set~$K_b = K_x = 35$.  For simplicity, we assume~$\Sigma$ is known in computation of the eigendecompositions. Given~$\{ X(t) \}_{0 \leq t \leq 1}$ for a given user-day, we generate event times according a chosen hazard function~$h (t; \theta)$.  To mimic our real data, we set
\[
h(t; \theta) = \exp \left( \theta_0 + \int_{0}^{\Delta} X(t-s) \beta(s) ds \right).
\]
We set~$\Delta$ to mimic a 30-minute window for a $12$-hour day.  We set~$\theta_0 = \log(5/1000)$ to set a baseline risk of approximately 5 events per day. We consider two choices of~$\beta(s)$: (1) $\beta_0 + \exp(- \beta_1 s) $, and (2) $\beta_1  * \sin \left( 2 \pi \frac{s}{\Delta} - \pi/2 \right)$.

Given each simulated user-day, we then
\begin{itemize}
\item Sample every 0.5 hours; then subsample at each subrate. Estimates are given by $\hat \beta_{i,0.5} (t)$ and $\hat \beta_{i,1} (t)$.
\item Bias is average over the estimates $\bar \hat \beta_{0.5} (t) = 100^{-1} \sum_{i=1}^100 \hat \beta_{i, 0.5} (t)$ compared to truth.
\item Take squared difference integrated to get variance.
\item To get the subsampling variance take $1/1000 \sum_{i=1}^1000 (\hat \beta_{i,0.5} (t) - \hat \beta_{i,1} (t))^2$.
\item Normalize everything by squared function!
\end{itemize}

We sample non-events at rates of 1 per 4/2/1/0.5 hours.

Since we are primarily interested in accuracy of the final estimates~$\hat \beta(s)$, we report the mean integrated squared error (MISE) of~$\hat \beta (s)$, defined as~$\frac{1}{1000} \sum_{j=1}^{1000} \int_{0}^{\Delta^\prime} ( \hat \beta_j (s) - \beta(s))^2 ds$ where~$\beta(s) \equiv 0$ for $s > \Delta$.  We define the MISE in this manner to account for settings where~$\Delta$ is unknown. Table~\ref{tab:mise} show the MISE decomposed into the variance and squared biase

\begin{table}[!th]
\begin{tabular}{c | p{2cm} p{2cm} p{2cm} p{2cm}}
& \multicolumn{4}{c}{Sampling rate} \\ \cline{2-5}
 & 0.5 & 1 & 2 & 4 \\ \hline
Subsampling variance & - & $9.0 \times 10^{-7}$ & $2.9 \times 10^{-6}$ & $5.8 \times 10^{-6}$  \\
Variance & $7.9 \times 10^{-3}$ & $9.2 \times 10^{-3}$ & $1.0 \times 10^{-2}$ & $1.30 \times 10^{-2}$ \\
Squared Bias & $9.8 \times 10^{-2}$ & $9.5 \times 10^{-2}$ & $9.3 \times 10^{-2}$ & $9.2 \times 10^{-2}$   \\
MISE & $1.1 \times 10^{-1}$ & $1.0 \times 10^{-1}$ & $1.0 \times 10^{-1}$ & $1.0 \times 10^{-1}$ \\ \hline
Avg. runtime (secs) & 317 & 177 & 104 & 68 \\ \hline
Subsampling variance & - & $2.9 \times 10^{-6}$ & $8.9 \times 10^{-6}$ & $2.0 \times 10^{-5}$  \\
Variance & $1.0 \times 10^{-2}$ & $1.3 \times 10^{-2}$ & $2.0 \times 10^{-2}$ & $3.1 \times 10^{-2}$ \\
Squared Bias & $6.9 \times 10^{-2}$ & $6.5 \times 10^{-2}$ & $6.0 \times 10^{-2}$ & $5.4 \times 10^{-2}$   \\
MISE & $7.9 \times 10^{-2}$ & $7.9 \times 10^{-2}$ & $8.0 \times 10^{-2}$ & $8.4 \times 10^{-2}$ \\ \hline
Avg. runtime (secs) & 36 & 23 & 15 & 12 \\ \hline
\end{tabular}
\caption{Mean-integrated squared error and type I error as a function of sampling rate for $\beta(s)$ given by (1) and (2) respectively.}
\label{tab:mise}
\end{table}

@@

\section{Discussion}
We abide by the comment of Robert Gentleman on big data: (1) make the data as small as possible as fast as possible. Statistics works!  With subsampling and proper re-weighting the analysis is converted to well-known territory. (2) Make the software and tools as fast and as maintanable as possible.  This is exactly teh benefit of subsampling.  we convert to penalized logistic regression on a smaller but still relatively efficient dataset. This means we can use all the existing tools!

Extensions to online and sampling over and over as we improve estimators.
